\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}

\setlength{\parskip}{0.25em}
\setlength{\parindent}{0em}

\begin{document}

\begin{center}
{ \LARGE Constitution of University of Ottawa Computer Science Club }

%  This should be the date of the last official change, meaning the day of the
%  vote for the change. TODO: review this policy
Last edited on 2023-05-11
\end{center}

\section*{Article 1 - Name}

The club’s official name will be ``University of Ottawa Computer Science Club''.
No other name will be used in the advertisement of representation of the club other than
the semi-abbreviated ``UOCSClub''.

\section*{Article 2 - Club Mandate}

% It should be general enough not to dictate us to perform specific activities,
% but to provide a guideline of the intention behind our events and activities.

\begin{enumerate}
    % Hackathons, workshops, and other events
    \item To host and promote social and competitive hacking challenges;
    % Managing our Slack channel, partially related to our social events, notifying
    % community of important developments in IT.
    \item To foster a hacker/coder community at the University of Ottawa;
    % 
    \item To provide coding sessions and technology workshops;
    \item To organize community-driven, system administration projects; and
    \item To be a forum for teaching and sharing IT knowledge.
\end{enumerate}

\section*{Article 3 - Membership}

% This part of the constitution is essential. We need to define what it means to
% be a member of CS Club.

An individual becomes a recognized \emph{member} of the club upon official registration with the club's executives.

The executives hold specific responsibilities and are mandated to oversee the membership of the club, including:

\begin{enumerate}
    \item Ensuring that membership is open, accessible, and free to all individuals;
    \item Ensuring that a minimum of 75\% of the club's members are registered students at the University of Ottawa; and
    \item Requiring that all executive members of the club are either registered students at the University of Ottawa or students who have graduated within the past academic year.
\end{enumerate}

\section*{Article 4 - Executives}

% TODO: We need to determine all of the official roles of the club.

% Need:
%  2 people to have signing authority
%  to define our council
%  (if CSSA thing happens) official representative
%  
The club shall be governed by 2 bodies: the Executive and the Council; defined here:
\begin{enumerate}
    \item The Executive committee shall consist of the following:
    \begin{enumerate}
        \item President;
        \item Vice President;
        \item Event Manager; and
        \item Treasurer.
    \end{enumerate}
    \item The Council shall consist of the Executive and members appointed by the Executive
\end{enumerate}

\section*{Article 5 - Responsibilities of the Executive}

All members of the Executive are members of the Council. The responsibilities of the
Executive is a superset of the responsibilities of the Council (as defined in Article 6)
and includes the additional responsibilities defined below:
\begin{enumerate}
    \item The President will:
    \begin{enumerate}
        \item Oversee the other members of the executive in fulfilling their
              responsibilities;
        \item Chair all meetings; and
        \item Have signing authority for the club.
    \end{enumerate}
    \item Vice-President will:
    \begin{enumerate}
        \item Assist the President in her / his duties;
        \item Assume all powers of the President in her / his absence;
        \item Be responsible for recording minutes of all meetings;
        \item Maintain an active record of club members; and
        \item Have signing authority for the club.
    \end{enumerate}
    \item The Event Manager will:
    \begin{enumerate}
        \item Assist in planning and overseeing any internal events held by the club.
    \end{enumerate}
    \item The Treasurer will:
    \begin{enumerate}
        \item Be responsible for overseeing all financial dealing of the club; and
        \item Keep complete records of all financial dealings of the club.
    \end{enumerate}
\end{enumerate}

\section*{Article 6 - Responsibilities of the Council}

The Council shall consist of Members appointed by the council and the executive
as defined in Article 8. All members of the council, including the executive, will be
responsible for the following:
\begin{enumerate}
    \item The members of the Council shall:
    \begin{enumerate}
        \item Prepare for and attend meetings of the council;
        \item Cast one (1) vote on all decisions of the meeting; and
        \item Perform duties appointed to them through meeting motions.
    \end{enumerate}
    \item The meetings of the Council shall be held to the standards of:
    \begin{enumerate}
        \item At least one (1) general meeting shall be held during every semester;
        \item Members will be informed of each of these meetings at least seven (7)
              days in advance;
        \item Quorum for a meeting of the Council is defined as one-half (1/2) of all appointed
              council members, including one (1) executive member;
        \item All decisions of the meeting require at least one-half plus one (1/2 + 1) of all votes
              to pass, with exceptions to decisions outlined within this document requiring a larger majority; and
        \item Minutes of the meetings be made available to members upon request to the Council.
    \end{enumerate}
\end{enumerate}

\section*{Article 7 - Elections}

% The election process is for the executive members.

% % This is the election process proposed by the SFUO
\begin{enumerate}
    \item Elections for the Executive will take place at the end of the winter semester.
    \begin{enumerate}
        \item No earlier than the last week of March.
        \item No later than the last week of April.
    \end{enumerate}
    % TODO: update working from election date until the next election
    \item The length of the mandate of executive members is from May 1st to
           April 30th of the following year.
    \item The executive positions will be open to members of the club who are
          students at the University of Ottawa or have graduated within 1 year
          of their election. % TODO: Add clause for expiry date
    \item Elections will start during a general meeting on a date determined by the Council. This date 
          must be set and conveyed to club members no
          less than two weeks in advance along with a call for members to submit
          their candidacy for the available executive positions. The campaigning period will start exactly one (1) week after the meeting is announced.
    \item Candidates for executive positions must convey their interest in a
          particular position by writing to the current executive before the
          campaigning period begins. Candidates also have the opportunity to submit
          a platform for the position, which will be circulated to club members
          at the beginning of the campaign period. In the event that there is nobody contesting any
         one of the executive positions in advance of the vote, interested persons
          may indicate their interest in the position at the elections meeting. % TODO: Add in a clause for remote voting
    \item During the election, each club member will have one (1) vote for each executive position. The voting period will last three (3) days, starting on the day of the election meeting.
    \item The vote shall be made via an online ballot.
    \item One council member will be appointed to hold a tie-breaking vote prior to the start of the election period. This member cannot run in the election. The tie-breaking vote shall be cast before the opening of the ballot box.
    \item The winner for each position will be the candidate with the greatest number of votes. In the case of a tie, the winner will be the recipient of the tie-breaking vote.
    \item Candidates may request for any individuals to scrutinize and count votes. If no request was made, the votes will be counted by the election officials. The election officials must be present during the vote counting either way.
\end{enumerate}

\section*{Article 8 - Amendments}

% CSSA version
% The procedure for amending this constitution shall be as follows:
% (i) A written report detailing the specific amendments shall be presented at a meeting
% of the Executive and the report shall be included in the minutes of the meeting;
% (ii) Following the meeting, there shall be a minimum of five (5) days and maximum
% fourteen (14) day reflection period during which the proposed amendments shall be publicly available;
% (iii) Following the reflection period, a meeting of the Executive shall be held and the
% proposed amendments will be voted on.
% 2. The date of the most recent ratification of this document shall be included at the
% beginning of this document.
% 3. All amendments to this document shall be annotated with the date of the ratification
% of the amendment.

% Rodgers version


% Abdulwahaab version
% Same as CSSA, just add if the change affects members of the club, then members can vote.
% If the changes don't affect the members of the clubs, then only the council votes.

% I believe this part is done.
\begin{enumerate}
    \item Amendments to the constitution must win a two-thirds (2/3) majority
          vote of the members present; and,
    \item An amendment to the constitution must be published to the members, which
          must be presented with a typed copy of the amendment as well as typed
          minutes from the meeting when the amendment
          was passed in order to prove that the amendment was passed.
\end{enumerate}

\section*{Article 9 - Impeachment}

\begin{enumerate}
    \item In the event that a club member engages in conduct that detrimentally impacts the welfare of the club and its members, they may be formally notified of impeachment.
    \item The individual facing impeachment shall be granted the opportunity to present a defence for their actions.
    \item Upon a two-thirds (2/3) majority vote of the members present, the impeached individual will be expelled from the club and forfeit all privileges associated with club membership.
\end{enumerate}

\section*{Article 10 - Executive Removal Process}

\begin{enumerate}
    \item In the event that the President of the University of Ottawa Computer Science Club engages in actions that are detrimental to the club's interests or fails to fulfill their responsibilities, the general members and executives may initiate a removal process. The process for the removal of the President shall follow the steps outlined below:
    
    \begin{enumerate}
        \item Notice of impeachment: Any member or executive can submit a written notice of impeachment to the Council, detailing the specific actions or behaviours that warrant removal. The notice of impeachment should be presented during a council meeting, and a copy of the notice shall be provided to the impeached President.
        
        \item Right to defend: The impeached President shall have the right to defend themselves during a special meeting called for this purpose. This meeting shall be held within a reasonable timeframe after the notice of impeachment has been submitted. The President may present their defence and provide evidence or witnesses in their favour.
        
        \item Voting for removal: After the President has presented their defence, a vote for removal shall be conducted. A two-thirds (2/3) majority vote of the members present at the meeting, including both general members and executives, shall be required to remove the President from their position.
        
        \item Loss of privileges: If the vote for removal is successful, the impeached President shall be immediately removed from their position and lose any privileges associated with the club's presidency. The President's duties and responsibilities shall be transferred to the Vice President or another designated executive member until a new President is elected.
    \end{enumerate}
    
    \item If any executive member, other than the President, engages in actions that are detrimental to the club's interests or fails to fulfill their responsibilities, the general members and other executives may initiate a removal process. The process for the removal of executives shall follow a similar procedure as outlined for the removal of the President, with necessary modifications to account for the specific executive position in question.
\end{enumerate}


% \section*{Article 10 - Finance}
% 
% \begin{enumerate}
%     \item The Executives will set a membership fee annually.
%     \item Additional fees may be collected for voluntary participation in
%           club events with external costs.
% \end{enumerate}

% \section*{Article 11 - Refund Policy}

% \begin{enumerate}
%     \item Initiating a membership fee refund policy for club members which is to include the
%           following format and minimum standard:
%     \begin{enumerate}
%         \item A member may apply to her club for a refund within one (1) month
%               of becoming a member of the club, or within one (1) week of the
%               club’s first official event if:
%         \begin{enumerate}
%             \item There has been a misinterpretation of the club’s mandate
%                   and proposed activities as specified by the member when
%                   signing onto the club.
%         \end{enumerate}
%         \item A member may only apply to her club for a refund after one (1)
%               month of signing up for membership for the club, or after one (1)
%               week of the club’s first official event for extenuating
%               circumstances.
%         \begin{enumerate}
%             \item Extenuating circumstances include, but are not limited to:
%                   Serious organizational issues with the club executive that
%                   led to a complete lack of communication with its members, or
%                   lack of programming as promoted to its members; and, any
%                   circumstances that seriously hamper the ability of the club
%                   members to enjoy their membership to the club.
%         \end{enumerate}
%     \end{enumerate}
%     % TODO: Review this refund policy
%     \item No refund shall be given for payments related to a member's voluntary participation
%         in a Club event with external costs.
% \end{enumerate}

\section*{Article 11 - Agency Clause}

% This would need to be updated to include CSSA or other overseeing organizations.
\begin{enumerate}
    \item The University of Ottawa Computer Science Club is not an agent of
          the University of Ottawa and its views and actions do not represent those
          of the University of Ottawa.
\end{enumerate}

% Include a discussion on club meetings

\end{document}

